---
title       : rCharts-Tutorial
subtitle    : 
author      : Vivek Patil
job         : Associate Professor of Marketing
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## rCharts

* Provides the ability, from within R, to create, customize and publish interactive JavaScript (JS) visualizations
* JS is a computer programming language that can create interactivity using web browsers
* There are many pre-written JS libraries that make the process of creating visualizations easier. 
* rCharts provides wrapper functions around few, but among the best, of these JS libraries to let an R user create interactive JS visualizations with little knowledge of JS. 

---

## JS libraries currently supported (as of 10/2014)

|   |   |
|---|---|
| [datatables](http://www.datatables.net/)  | Interactive tables  |
| [dimple](http://dimplejs.org/)  | d3-based - plenty of options  |
| [highcharts](http://www.highcharts.com/)  |Free only for non-commercial purposes, plenty of options   |
| [morris](http://morrisjs.github.io/morris.js/)  | currently - line, bar, area and donut charts  |
| [nvd3](http://nvd3.org/)  | d3-based - plenty of options   |
| [polychart](https://www.polychart.com/js/)  | Free for non-commercial purposes only  |
| [rickshaw](http://code.shutterstock.com/rickshaw/)  |d3-based, time-series specialty but many others |
| [timeline](http://timeline.knightlab.com/)  | Interactive timelines  |
| [uvcharts](http://imaginea.github.io/uvCharts/)  | d3-based   |
| [xcharts](http://tenxer.github.io/xcharts/)  |  Examples of line and bar charts and custom modifications of these  |

* datamaps, leaflet, and crosslet are 3 others that have been wrapped up in another package called rMaps (for a different day)

---

## Challenge in learning

* Different JS libraries developed independently - consistency in calls from within rCharts can at times be a challenge
* Customization to specific needs may require writing JS code

### Learning approach

* Follow examples, questions, and issues presented in different venues
* Stick to rCharts interface to few JS libraries which provide many choices of charts and become very familiar with them. Don't forget that few have restrictions for commercial use.

---

## Few Resources

* [rCharts' website](http://rcharts.io/) and its [gallery of submissions from different people](http://rcharts.io/gallery/).
* [rCharts on GitHub and the issues posted there](https://github.com/ramnathv/rCharts) and on [stackoverflow](http://stackoverflow.com/search?q=rCharts)
* [A work-in-progress documentation](http://timelyportfolio.github.io/docs/_build/html/)
* [rCharts-related posts from different bloggers via R-bloggers](http://www.r-bloggers.com/?s=rCharts)


|Code examples for different libraries   |   |
|---|---|
| [datatables](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/datatables/examples.R)  | [polychart](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/polycharts/examples.R)    |
| [dimple](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/dimple/examples.R)  | [rickshaw](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/rickshaw/examples.R)  |
| [highcharts](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/highcharts/examples.R)  | [timeline](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/timeline/examples.R)  |
| [morris](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/morris/examples.R)  | [uvcharts](http://imaginea.github.io/uvCharts/)  |
| [nvd3](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/nvd3/examples.R)  | [xcharts](https://github.com/ramnathv/rCharts/blob/master/inst/libraries/xcharts/examples.R)   |

---

## Datatable

```{r setup, message = F, echo = F, cache = F}
require(rCharts)
knitr::opts_chunk$set(results = 'asis', cache=FALSE, message=FALSE, warning=FALSE)
```

```{r}
dt <- dTable(
  iris,
  sPaginationType= "full_numbers"
)
dt$save('datatable.html',cdn=TRUE)
```
<iframe src="datatable.html"></iframe>

---
## Dimple - Verticle stacked & Grouped Bar Chart (the documentation and examples are most extensive)

```{r}
data <- read.delim(
  "http://pmsi-alignalytics.github.io/dimple/data/example_data.tsv"
)
colnames(data) <- gsub("[.]","",colnames(data))
d1 <- dPlot(
  x = c("PriceTier","Channel"),
  y = "UnitSales",
  groups = "Owner",
  data = data,
  type = "bar"
)
d1$legend(
  x = 200,
  y = 10,
  width = 400,
  height = 20,
  horizontalAlign = "right"
)
d1$save('dimple.html',cdn=TRUE)
```

---

## Dimple: Verticle stacked & Grouped Bar Chart (the documentation and examples are most extensive)

<iframe src="dimple.html"></iframe>

---

## Highchart: Bubble Chart with Zoom and Export (plenty of other cool ones)

```{r}
h1 <- hPlot(Pulse ~ Height, data = MASS::survey, type = "bubble", title = "Zoom demo", subtitle = "bubble chart", size = "Age", group = "Exer")
h1$chart(zoomType = "xy")
h1$exporting(enabled = T)
h1$show("inline")
```

---
# Morris - line chart

```{r}
data(economics, package = 'ggplot2')
dat = transform(economics, date = as.character(date))
m1 <- mPlot(x = "date", y = list("psavert", "uempmed"), data = dat, type = 'Line',
            pointSize = 0, lineWidth = 1)
m1$set(xLabelFormat = "#! function (x) { 
  return x.toString(); } 
!#")
m1$save("morris.html",cdn=TRUE)
```
<iframe src="morris.html"></iframe>

---
## nvd3: Grouped/Stacked Bar Chart (plenty of cool examples - check out the line with focus)

```{r}
hair_eye = as.data.frame(HairEyeColor)
n1 <- nPlot(Freq ~ Hair, group = 'Eye', data = subset(hair_eye, Sex == "Female"), type = 'multiBarChart')
n1$chart(color = c('brown', 'blue', '#594c26', 'green'))
n1$save('nvd3.html',cdn=TRUE)
```
<iframe src="nvd3.html"></iframe>

--- 
## polycharts: facetted bar plot (also check out their heatmap)

```{r}
names(iris) = gsub("\\.", "", names(iris))
p1 <- rPlot(SepalLength ~ SepalWidth | Species, data = iris, color = 'Species', type = 'point')
p1$save("polychart.html",cdn=TRUE)
```
<iframe src="polychart.html"></iframe>

---
## Rickshaw: Area chart

```{r}
uspexp <- reshape2::melt(USPersonalExpenditure)
names(uspexp) <- c('category', 'year', 'expenditure')
to_jsdate <- function(date_){
  val = as.POSIXct(as.Date(date_), origin="1970-01-01")
  as.numeric(val)
}
uspexp <- transform(uspexp, year = to_jsdate(as.Date(paste(year, '01', '01', sep = '-'))))
r1 <- Rickshaw$new()
r1$layer(expenditure ~ year, group = 'category', data = uspexp, type = 'area')
r1$yAxis(orientation = 'left')
r1$xAxis(type = 'Time')
r1$set(width = 700, height = 400)
r1$save("rickshaw.html",cdn=TRUE)
```

---
## Rickshaw: Area chart

<iframe src="rickshaw.html"></iframe>

---

## UV Charts

```{r}
u1=uPlot(x = 'Sex', y = 'Freq', data = as.data.frame(HairEyeColor), group = 'Hair', type = 'Bar')
u1$save("uv.html",cdn=TRUE)
```
<iframe src="uv.html"></iframe>

---

## Displaying/Distributing them

* `chartname$save('chartname.html')` and then use iframe 
* `chartname$publish('chartname.html')` # need a GitHub account

(See [rCharts Issue# 373 for other options](https://github.com/ramnathv/rCharts/issues/373))

---

## Few Resources, again

* [rCharts' website](http://rcharts.io/) and its [gallery of submissions from different people](http://rcharts.io/gallery/).
* [rCharts on GitHub and the issues posted there](https://github.com/ramnathv/rCharts) and on [stackoverflow](http://stackoverflow.com/search?q=rCharts)
* [A work-in-progress documentation](http://timelyportfolio.github.io/docs/_build/html/)
* [rCharts-related posts from different bloggers via R-bloggers](http://www.r-bloggers.com/?s=rCharts)

* [One of my posts on different charts rCharts provides for different questions](http://patilv.com/Replication-of-few-graphs-charts-in-base-R-ggplot2-and-rCharts-part-3-rCharts/)

Code for this presentation can be found at: [https://github.com/patilv/rCharts-tutorial](https://github.com/patilv/rCharts-tutorial)
